---
title: "Simulation exercise using the exponential function"
author: "Rohil Narula"
date: "20 July 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulation exercise

## Overview

This simulation exercise will draw $40$ variables from a random exponential distributtion of data with $\lambda = 0.2$. A mean distribution of $1000$ simulations will be sampled and the mean and variance of this distribution, calculated. Finally, a simple histogram plot will be used to visualize the distribution.

## Simulations

In this section, the initial variables are set, then each sample of the simulations will be stored as a row in a matrix. As we have $1000$ simulations, the matrix will by default contain $1000$ rows (as well as $40$ columns).

The initial variables are set first.

```{r init}
n <- 40
lambda <- 0.2
draws <- 1000
```

The matrix of the random draws is calculated using the standard `R` function `rexp()`.

```{r matrix}
simulations <- matrix(rexp(n * draws, lambda), nrow = draws)
str(simulations)
```